Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-03-23 src/ms.tex
paper edits

\section{The Gaia Color–Period Plane}\label{app:gaia}
\edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high precision photometric colors provided by the Gaia mission \citep{Gaia2016}. We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3}. We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves. For this exercise, we used the empirically calibrated \taucz relation of \citet{Corsaro2021}, who presented \taucz as a quadratic function of Gaia $G_{BP}-G_{RP}$ color using the Kepler LEGACY asteroseismic sample as calibrators. We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded. In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=0.5$ which is approximately the solar value ($\mathrm{Ro_\odot}=0.496$) on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = 0.5 \approx \mathrm{Ro}_\odot$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$. The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data. Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above. The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
    \label{fig:gaia-mcquillan}
\end{figure}

>>>>

\section{The Gaia Color–Period Plane}\label{app:gaia}
\edit1{Given that the long-period pile-up was previously obscured by imprecise \teff measurements, we explored whether the feature could be recovered using the high precision photometric colors provided by the Gaia mission \citep{Gaia2016}. We cross-matched the \citet{McQuillan2014} sample with Gaia EDR3 \citep{GaiaEDR3}. We then compared the ($G_{BP}-G_{RP}$)–\prot distribution with constant Rossby curves. For this exercise, we used the empirically calibrated \taucz relation of \citet{Corsaro2021}, who presented \taucz as a quadratic function of Gaia $G_{BP}-G_{RP}$ color using the Kepler LEGACY asteroseismic sample as calibrators. We found that the long-period pile-up is clearly visible in the ($G_{BP}-G_{RP}$)–\prot plane once stars with high reddening ($A_V>0.2$) are excluded. In order to match a curve of constant Rossby number, given by $\mathrm{Ro}=\mathrm{Ro_\odot}=0.496$ on the \citet{Corsaro2021} scale, we found that a -0.1 mag shift to the Gaia colors of the data was required.}

\edit1{We note that the morphology of the color–period distribution appears to be different between the Gaia–McQuillan and Gaia–Santos catalogs, with the Gaia–Santos distribution presenting a break near $G_{BP}-G_{RP}= 0.7$. This break is not apparent in the Gaia–McQuillan sample. Similar behavior is seen in \teff–\prot plane, as shown in Figure~\ref{fig:xmatch}. At present, we do not have a satisfactory explanation for this behavior.}

\begin{figure}
    \includegraphics[width=\linewidth]{./figures/gaia-mcquillan.pdf}
    \caption{The color–\prot plane for the \citet{McQuillan2014} sample in relation to a curve of constant Rossby number (orange dashed curve, $\mathrm{Ro} = \mathrm{Ro}_\odot = 0.496$) from the empirical \taucz calibration of \citet{Corsaro2021}, which is valid in the color range $0.55 < (G_\mathrm{BP}−G_\mathrm{RP}) < 0.97$. The top rows show 2–d histograms of the probability density for the entire sample (left), a subsample with low reddening (middle), and the same low reddening subsample with a -0.1 mag color shift applied to the data. Each panel in the bottom row shows a Gaussian kernel density estimation of the respective panel above. The orange point in each panel represents the position of the Sun, using the estimated Gaia color of \citet{Casagrande2018}.}
    \label{fig:gaia-mcquillan}
\end{figure}
